Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2022-08-11 manuscript/Chapter3-manuscript.Rmd
updt MS. updt validation. updt loo 02A1. updt bib

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@limaPuttingPredatorsBack2002;@schreiberCommunityEffectsPhenotypic2011;@pettorelliChapterTwoIndividual2015;@Michalko.Pekar2016]. Indeed, predator populations are often composed of assemblages of individuals specializing in different foraging strategies or resources [@estesIndividualVariationPrey2003;@tinkerFoodLimitationLeads2008;@kernaleguenDeterminantsIndividualForaging2015;@Phillips.etal2017]. It is now well established that individual foraging specialization cannot be attributed exclusively to intrinsic factors such as sexual, morphological, or age-related differences [@bolnickEcologyIndividualsIncidence2003;@bolnickWhyIntraspecificTrait2011]. A growing body of evidence suggests that ecological interactions, such as predator-prey interactions, may instead act as the major drivers of individual foraging specialization [@araujoEcologicalCausesIndividual2011;@toscanoPersonalityForagingBehavior2016]. For instance, predators species often use hunting techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@estesIndividualVariationPrey2003;@Woo.etal2008;@courbinShorttermPreyField2018], and their capacity to learn how to effectively use them is essential to maintain foraging success. Yet, we have very limited knowledge on how predator foraging specialization develops through experience and interactions with prey, and the resulting ecological consequences of such among individual behavioural differences for predator-prey interactions.

Learning how to hunt prey is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@heithausFeedingStrategiesTactics2018]. Theory predicts that learning may promote foraging specialization via past experiences, extensive practice, as well as a limited ability to retain multiple complex hunting skills [@Tinker.etal2009;@Dukas2019]. From this perspective, learning is an optimizing process that enables foraging specialization. Empirical studies show that learning allows predators to optimize the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@edwardsRoleExperienceDevelopment1994;@morseEffectExperienceHunting2000;@reidLearningHuntRole2010;@wilson-rankinLevelExperienceModulates2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics when prey are scarce or highly unpredictable [@estesIndividualVariationPrey2003;@Mery.Burns2010]. An alternative mechanism is that learning may offset the costs/risks of switching foraging tactics as individuals gain experience and information on their prey or their environment, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013]. In this case, switching between foraging tactics would be advantageous when resources are variable, as behavioural flexibility is often key for survival in novel environments [reviewed in @Snell-Rood2013]. If learning is the mechanism leading to individual specialization, then it wouldn't explain why field studies often find that foraging specialists and generalists have similar hunting success irrespective of their age [@Woo.etal2008]. *(ne peut pas être séparé de l'interaction car c'ets ce qui drive l'apprentissage)*

@@

@@

(travailler ceci) An important question that remains unanswered is how can generalist and specialist individuals coexist within predator populations. If foraging and resource specialisation can help maintain community stability by promoting species coexistence [sources: bolnick, araujo, wolf, etc], it is crucial that we develop studies that integrate the causes and consequences of these two individual components. Several studies have shown that individual specialists can outperform. *[ici je crois que je vais changer un peu cela. C'est surtout des idées. Ici on veut souligner l'importance des deux mais rappeler comme d'habitude que c'est difficile à mesurer à travers le temps blabla. J'introduis ensuite le jeu vidéo]* *vidfeogames offer a good alternative to lab studies because proche de wild = observational*
*(idéee: par contre, décomposer l'âge de l'expérience est difficile, et particulièrement chez des espèces sur le terrain puisque ça peut impliquer des techniques invasives)*

>>>>

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@limaPuttingPredatorsBack2002;@schreiberCommunityEffectsPhenotypic2011;@pettorelliChapterTwoIndividual2015;@Michalko.Pekar2016]. Indeed, predator populations are often composed of assemblages of individuals specializing in different foraging strategies or resources irrespective of sexual, morphological, or age-related differences [@estesIndividualVariationPrey2003;@tinkerFoodLimitationLeads2008;@kernaleguenDeterminantsIndividualForaging2015;@Phillips.etal2017] (FIND SOME TERRESTRIAL ANIMAL REFS). A growing body of evidence suggests that ecological interactions, such as predator-prey interactions, can be major drivers of such individual foraging specialization [@araujoEcologicalCausesIndividual2011;@toscanoPersonalityForagingBehavior2016]. When hunting, predators often use hunting techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@estesIndividualVariationPrey2003;@Woo.etal2008;@courbinShorttermPreyField2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. expertise). While the development of hunting expertise may be essential to maintain foraging success, we have have limited evidence for its role in predator foraging specialization, and the ecological/fitness consequences of such among individual behavioural differences for predator-prey interactions.

The integration of individual behavioral variation into studies of predator-prey interactions has been championned for some time, and a series of empirical papers have revelealed important insights on its consequences for functional responses, habitat use, and prey choice. However, a recurring challenge impeding empirical studies of predator-prey behavioral interactions is the difficulty of collecting data simultaneously on both the predator and prey. We have recently demonstrated that using an online predator-prey videogame can help uncover subtle details about the mechanisms shaping these behavioral interactions [@FraserFranco].

*ici, mettre le paragraphe sur la variation individuelle, et ensuite expliquer ce qu'on cherche à comprendre dans la littérature écologique comportementale*

The development of foraging strategies is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@heithausFeedingStrategiesTactics2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. [Dukas @Dukas2019] defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise is an optimizing process that promotes foraging specialization. Empirical studies show that expertise allows predators to optimize the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@edwardsRoleExperienceDevelopment1994;@morseEffectExperienceHunting2000;@reidLearningHuntRole2010;@wilson-rankinLevelExperienceModulates2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics when prey are scarce or highly unpredictable [@estesIndividualVariationPrey2003;@Mery.Burns2010]. An alternative mechanism is that expertise may offset the costs/risks of switching foraging tactics as individuals gain experience and information on their prey or their environment, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013]. In this case, switching between foraging tactics would be advantageous when resources are variable, as behavioural flexibility is often key for survival in novel environments [reviewed in @Snell-Rood2013] (FIND OTHER REFS TO SUPPORT THIS MAYBE DAN SOL?).

@@

*ou juste dire que pour comprendre le role de l'expertise, il faut s'attarder à l'étudier avec la specialisation et voir les consequences pour la fitness ou issue des interactions predateur proie. ca introduit mieux je crois. dire genre que une théorie prédit que spécialisation est avantage quand proies peu variable, et ensuite ça lie bien l'hypothèse de variabilité que je décris en bas*

@@

 *Difficile de mesurer à travers le temps blabla. J'introduis ensuite le jeu vidéo* *vidfeogames offer a good alternative to lab studies because proche de wild = observational*
*(idéee: par contre, décomposer l'âge de l'expérience est difficile, et particulièrement chez des espèces sur le terrain puisque ça peut impliquer des techniques invasives)*
